SPECS-decadal-MOS

### 2017-07-07 SPECS/specs-decade-mos-all.Rmd
SPECS decadal multi-model analysis

@@

downloadfromunican <- FALSE    # Preparations - get the data
downscaledreanalysis <- FALSE  # SM: Assessment of large-small scale link
looexperiments <- FALSE        # Cross-validation MOS on forecasts
diagnoseFCST <- FALSE          # Show correlation & common EOFs of reanalysis+fcsts

source("/home/rasmusb/disk1/SPECS/R/udg2esd_met.R") ## Upload udg2esd function

@@

    ## The index is set by the year and the calendar month/day reflects
    ## member number
    if (d[2]<=12) index(Y1) <- as.Date(paste(year(Y1),im,'01',sep='-'),
                                       origin = as.Date("1950-01-01")) else
                  index(Y1) <- as.Date(paste(year(Y1),'01-01',sep='-')+im,
                                       origin = as.Date("1950-01-01"))
    if (verbose) print(index(Y1))

@@

  ## repeat the observations with corrsponding time stamps as the ensemble members of the forecast
  if (verbose) print('repeat predictand')
  for (i in 1:n) {
    if (verbose) print(paste('predictand',i))
    if (attr(X,'member_by')=='month')
      index(y) <- as.Date(paste(year(y),i,'01',sep='-'),
                          origin = as.Date("1950-01-01")) else
      index(y) <- as.Date(paste(year(y),'01-01',sep='-')+i,
                          origin = as.Date("1950-01-01"))
    if (verbose) print(index(y))
    if (i==1) Y <- y else Y <- c(Y,y)
  }
  ## Copy all attributes etc.
  Y <- attrcp(Y0,Y)
  class(Y) <- c('field','day','zoo')
  attr(Y,'dimensions') <- c(length(lon(Y0)),length(lat(Y0)),length(index(Y)))
  if (verbose) print('predictand sucessfully processed')
  ## The experiment
  xeof <- EOF(mask(X,land=TRUE))
  yrs <- year(X)[is.element(year(X),year(Y))]
  yrs <- yrs[!duplicated(yrs)]
  if (verbose) {print(yrs); print('LOO x-val...')}
  ## LOO
  for (it in 1:length(yrs)) {
    ## set index to exclude all members for the same evaluation interval
    iloo <- !is.element(year(Y),yrs[it])
    yloo <- subset(Y,it=iloo)
    if (verbose) {
      print(paste('LOO: it=',it,'years:'))
      print(table(year(yloo)))
    }
    class(yloo) <- class(X)
    yeof <- EOF(yloo,n=4)
    ## Some hacking needed to ensure that subset doesn't change the index type to year.
    class(yeof)[3] <- 'day';  class(xeof)[3] <- 'day';
    ds.loo <- DS(yeof,xeof,verbose=verbose)
    ## Extract the forecast for the evaluation interval:
    if (verbose) print('reconstruct field')
    jloo <- is.element(year(xeof),yrs[it])
    xval <- subset(as.field(predict(ds.loo,newdata=xeof)),it=jloo)
    ## combine each LOO result into one series:
    if (it==1) XVAL <- xval else
               XVAL <- c(XVAL,xval)
  }
  if (verbose) print('LOO completed')
  XVAL <- attrcp(Y0,XVAL); class(XVAL) <- class(Y)
  attr(XVAL,'dimensions') <- c(attr(Y,'dimensions')[1:2],length(index(XVAL)))
  ## Plot the correlation map between LOO results and original:
  rmap <- corfield(aggregate(XVAL,year,'mean'),aggregate(Y,year,'mean'),plot=FALSE)
  map(rmap,colbar=list(breaks=seq(-1,1,by=0.1),pal='t2m',rev=TRUE),type='fill',new=FALSE)
  invisible(XVAL)
}

@@

  X <- Xpredictor(x,m=l,it=it,is=is,mask=mask,verbose=verbose)
  n <- attr(X,'n_members')
  if (attr(X,'member_by')=='month') {
    if (verbose) print('separate ensemble members by calendar month')
    m <- as.integer(rownames(table(month(X))))
    j <- 1
    for (i in m) {
      y <- pentad(subset(X,it=month.abb[i]),l=l)
      index(y) <- as.Date(paste(year(y),i,'01',sep='-'))
      if (j==1) Y <- zoo(y) else
                Y <- c(Y,zoo(y))
      j <- j+1
    }
  }
  Y <- attrcp(X,Y)
  class(Y) <- class(X)
  attr(Y,'dimnames') <- NULL
  invisible(Y)
}

@@

```{r compare}
comparemannual <- function(tas_DECA,reanalysis='~/Downloads/air.mon.mean.nc',mask=TRUE,l= 3,eofs=NULL) {

@@

```{r diagnose}
diagnoseMOS <- function(results,y,l=1) {
  print(paste('diagnoseMOS',results))
  load(results)  
  eval(parse(text=paste('zmos <-',substr(results,14,nchar(results)-4))))
  attr(zmos,'variable') <- varid(y)
  print(range(year(zmos)))
  y <- pentad(y,it0=min(year(zmos)),l=l)
  print(index(y))
  print(index(zmos))
  print('ensemble mean: aggregate(.)')
  ## Different ensemble emmbers are represented as calendar months
  Z <- aggregate(zmos,year,'mean')
  index(Z) <- year(Z)
  print('correlation: ensemble mean FCST & reanalysis')
  r <- corfield(Z,y,plot=FALSE)
  map(r,colbar=list(breaks=seq(-1,1,by=0.1),pal='t2m',rev=TRUE),type='fill',new=FALSE)
  figlab('EOBS+ensemble mean MOS',ypos=0.999)

@@

print('Get the predictands: mu & fw from EOBS')
if (!file.exists('~/data/mu_0.25deg_reg.nc'))
  mu.eobs <-  wetdaymean('~/data/data.ECAD/rr_0.25deg_reg.nc') else
  mu.eobs <- retrieve('~/data/mu_0.25deg_reg.nc')
fw.eobs <- retrieve('~/data/fw_0.25deg_reg.nc')
attr(mu.eobs,'variable') <- 'mu'
attr(fw.eobs,'variable') <- 'fw'
attr(fw.eobs,'unit') <- 'fraction'
index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))

@@

## Test to examine the link between large-scale temperature and wet-day mean precip
## Derive the saturation vapour pressure from temperature
## retrieve did not get the right time stamp:
  es <- annual(C.C.eq(retrieve('~/Downloads/air.mon.mean.nc',lon=c(-60,40),lat=c(20,70))))
  es.eof <- EOF(mask(es,land=TRUE))

## Also get the mean sea-level pressure for the wet-day frequency 
  slp <- annual(retrieve('slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
  slp.eof <- EOF(slp)

@@

  class(mu.eof) <- class(es.eof)
  mu.ds <- DS(mu.eof,es.eof)
  class(fw.eof) <- class(slp.eof)
  fw.ds <- DS(fw.eof,slp.eof)

  plot(mu.ds,colbar1=list(breaks=seq(-0.025,0.025,by=0.005),rev=TRUE),
       colbar2=list(breaks=seq(-350,350,by=25)),type='fill',new=FALSE,)
  figlab('ESD(NCEP/NCAR)',ypos=0.999)

@@

print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
load('~/data/gfdl_DECA.rda')
psl_DECA.gfdl <- psl_DECA; rm('psl_DECA'); tas_DECA.gfdl <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
#> dim(psl_DECA.gfdl$Data)
#[1]   9  10 480  41  45
print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
load('~/data/hadcm3_DECA.rda')
psl_DECA.hadcm3 <- psl_DECA; rm('psl_DECA'); tas_DECA.hadcm3 <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
#> dim(psl_DECA.hadcm3$Data)
#[1]   6  10 480  33  30
load('~/data/miroc5_DECA.rda')
psl_DECA.miroc5 <- psl_DECA; rm('psl_DECA'); tas_DECA.miroc5 <- tas_DECA; rm('tas_DECA'); gc(reset=TRUE)
#> dim(psl_DECA.miroc5$Data)
#[1]   6   6 480  59  79
rm("psl_DECA_annual","tas_DECA_annual"); gc(reset=TRUE)
## Need to regrid (regrid.matrix) and combine the different forecasts into one object:
```

@@

if (looexperiments) {
## Carry out a LOO cross-validation experiment
  print('Check results for LOO experiment - if absent do it')
  if (!file.exists('specsdecadal.z.fw.y1.rda')) {
    print('do specsdecadal.z.fw.y1.rda')
    z.fw.y1 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.rda',z.fw.y1)
    rm("z.fw.y1"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.rda')) {
    print('do specsdecadal.z.mu.y1.rda')
    z.mu.y1 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.rda',z.mu.y1)
    rm("z.mu.y1"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.3.rda')) {
    print('do specsdecadal.z.fw.y1.3.rda')
    z.fw.y1.3 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:3,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.3.rda',z.fw.y1.3)
    rm("z.fw.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.3.rda')) {
    print('do specsdecadal.z.mu.y1.3.rda')
    z.mu.y1.3 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:3,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.3.rda',z.mu.y1.3)
    rm("z.mu.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.5.rda')) {
    print('do specsdecadal.z.fw.y1.5.rda')
    z.fw.y1.5 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:5,it=NULL,is=NULL,mask=FALSE)
    save(file='specsdecadal.z.fw.y1.5.rda',z.fw.y1.5)
    rm("z.fw.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.mu.y1.5.rda')) {
    print('do specsdecadal.z.mu.y1.5.rda')
    z.mu.y1.5 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:5,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.5.rda',z.mu.y1.5)
    rm("z.mu.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists('specsdecadal.z.fw.y1.9.rda')) {
    print('do specsdecadal.z.fw.y1.9.rda')
     z.fw.y1.9 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:9,it=NULL,is=NULL,mask=FALSE)
     save(file='specsdecadal.z.fw.y1.9.rda',z.fw.y1.9)
     rm("z.fw.y1.9"); gc(reset=TRUE)
   }
  if (!file.exists('specsdecadal.z.mu.y1.9.rda')) {
    print('do specsdecadal.z.mu.y1.9.rda')
    z.mu.y1.9 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:9,it=NULL,is=NULL,mask=TRUE)
    save(file='specsdecadal.z.mu.y1.9.rda',z.mu.y1.9)
    rm("z.mu.y1.9"); gc(reset=TRUE)
  }

## Diagnostics:
  print('diagnose the MOS results')
  load('specsdecadal.z.mu.y1.3.rda')
  it0 <- range(year(z.mu.y1.3))
  print(paste('time period',paste(it0,collapse='-')))
  score.mu.y1 <- diagnoseMOS('specsdecadal.z.mu.y1.rda',mu.eobs)
  score.fw.y1 <- diagnoseMOS('specsdecadal.z.fw.y1.rda',fw.eobs)
  score.mu.y1.3 <- diagnoseMOS('specsdecadal.z.mu.y1.3.rda',mu.eobs,l=3)
  score.fw.y1.3 <- diagnoseMOS('specsdecadal.z.fw.y1.3.rda',fw.eobs,l=3)
  score.mu.y1.5 <- diagnoseMOS('specsdecadal.z.mu.y1.5.rda',mu.eobs,l=5)
  score.fw.y1.5 <- diagnoseMOS('specsdecadal.z.fw.y1.5.rda',fw.eobs,l=5)
  score.mu.y1.9 <- diagnoseMOS('specsdecadal.z.mu.y1.9.rda',mu.eobs,l=9)
  score.fw.y1.9 <- diagnoseMOS('specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)
  while (dev.cur()>1) dev.off()
## Repeat and make a table of the results...

@@

if (diagnoseFCST) {
  comparemannual(tas_DECA,mask=TRUE,l= 1)
  comparemannual(tas_DECA,mask=TRUE,l= 2)
  comparemannual(tas_DECA,mask=TRUE,l= 4)
  comparemannual(tas_DECA,mask=TRUE,l= 5)
  comparemannual(tas_DECA,mask=TRUE,l= 6)
  comparemannual(tas_DECA,mask=TRUE,l= 7)
  comparemannual(tas_DECA,mask=TRUE,l= 8)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 1)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 2)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 3)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 4)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 5)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 6)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 7)
  comparemannual(psl_DECA,reanalysis='slp.mon.mean.nc',mask=FALSE,l= 8)
  rmap.1.10.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:10)
  rmap.1.10.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:10)
}
```

@@

```{r t2m scales}
## Scaling dependency: time - space
t2m <- retrieve('~/Downloads/air.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
t2m <- anomaly(t2m)
zst <- scaledependency(t2m)
#zst[,1] <- NA
d <- dim(zst)

@@

## Repeat for SLP
slp <- retrieve('~/data/ncep/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
slp <- anomaly(slp)
yst <- scaledependency(slp)
image(attr(yst,'timescales'),attr(yst,'gridboxes'),t(t(yst)/colMeans(yst)),
      main='Mean sea-level pressure: spectral power',
      xlab='Temporal scale (years)',ylab='spatial scale (grid boxes)')
contour(attr(yst,'timescales'),attr(yst,'gridboxes'),yst,add=TRUE)

>>>>

Use `purl()` to generate a plain R-script.

@@

downloadfromunican <- FALSE    # Preparations - get the data
downscaledreanalysis <- TRUE  # SM: Assessment of large-small scale link
looexperiments <- TRUE        # Cross-validation MOS on forecasts
diagnoseFCST <- TRUE          # Show correlation & common EOFs of reanalysis+fcsts

#source("/home/rasmusb/disk1/SPECS/R/udg2esd_met.R") ## Upload udg2esd function

@@

    ## The index is set by the year and the calendar month/day reflects
    ## member number
    if (verbose) print(year(Y1))
    if (d[2]<=12) index(Y1) <- as.Date(paste(year(Y1),im,'01',sep='-'),
                                       origin = as.Date("1950-01-01")) else
                  index(Y1) <- as.Date(as.Date(paste(year(Y1),'01-01',sep='-'))+im,
                                       origin = as.Date("1950-01-01"))
    if (verbose) print(index(Y1))

@@

  ## repeat the observations with corrsponding time stamps as the ensemble members of the forecast
  if (verbose) print('repeat predictand')
  for (i in 1:n) {
    if (verbose) print(paste('predictand',i))
    if (attr(X,'member_by')=='month')
      index(y) <- as.Date(paste(year(y),i,'01',sep='-'),
                          origin = as.Date("1950-01-01")) else
      index(y) <- as.Date(as.Date(paste(year(y),'01-01',sep='-'))+i,
                          origin = as.Date("1950-01-01"))
    if (verbose) print(index(y))
    if (i==1) Y <- y else Y <- c(Y,y)
  }
  ## Copy all attributes etc.
  Y <- attrcp(Y0,Y)
  class(Y) <- c('field','day','zoo')
  attr(Y,'dimensions') <- c(length(lon(Y0)),length(lat(Y0)),length(index(Y)))
  if (verbose) print('predictand sucessfully processed')
  ## The experiment
  xeof <- EOF(mask(X,land=TRUE))
  yrs <- year(X)[is.element(year(X),year(Y))]
  yrs <- yrs[!duplicated(yrs)]
  if (verbose) {print(yrs); print('LOO x-val...')}
  ## LOO
  for (it in 1:length(yrs)) {
    ## set index to exclude all members for the same evaluation interval
    iloo <- !is.element(year(Y),yrs[it])
    yloo <- subset(Y,it=iloo)
    if (verbose) {
      print(paste('LOO: it=',it,'years:'))
      print(table(year(yloo)))
    }
    class(yloo) <- class(X)
    yeof <- EOF(yloo,n=4)
    ## Some hacking needed to ensure that subset doesn't change the index type to year.
    class(yeof)[3] <- 'day';  class(xeof)[3] <- 'day';
    ds.loo <- DS(yeof,xeof,verbose=verbose)
    ## Extract the forecast for the evaluation interval:
    if (verbose) print('reconstruct field')
    jloo <- is.element(year(xeof),yrs[it])
    xval <- subset(as.field(predict(ds.loo,newdata=xeof)),it=jloo)
    ## combine each LOO result into one series:
    if (it==1) XVAL <- xval else
               XVAL <- c(XVAL,xval)
  }
  if (verbose) print('LOO completed')
  XVAL <- attrcp(Y0,XVAL); class(XVAL) <- class(Y)
  attr(XVAL,'dimensions') <- c(attr(Y,'dimensions')[1:2],length(index(XVAL)))
  ## Plot the correlation map between LOO results and original:
  rmap <- corfield(aggregate(XVAL,year,'mean'),aggregate(Y,year,'mean'),plot=FALSE)
  map(rmap,colbar=list(breaks=seq(-1,1,by=0.1),pal='t2m',rev=TRUE),type='fill',new=FALSE)
  invisible(XVAL)
}

@@

  X <- Xpredictor(x,m=l,it=it,is=is,mask=mask,verbose=verbose)
  n <- attr(X,'n_members')
  if (attr(X,'member_by')=='month') {
    if (verbose) print('separate ensemble members by calendar month')
    m <- as.integer(rownames(table(month(X))))
    j <- 1
    for (i in m) {
      y <- pentad(subset(X,it=month.abb[i]),l=l)
      index(y) <- as.Date(paste(year(y),i,'01',sep='-'))
      if (j==1) Y <- zoo(y) else
                Y <- c(Y,zoo(y))
      j <- j+1
    }
  } else {
    jday <- julian(index(x))-julian(as.Date(paste(year(x),'01-01',sep='-')))
    m <- as.integer(rownames(table(jday)))
    if (verbose) print(paste('separate ensemble members by day 1 --',m))
    j <- 1
    for (i in m) {
      itday <- is.element(jday,i)
      if (verbose) print(c(i,j,sum(itday)))
      y <- pentad(subset(X,it=itday),l=l)
      index(y) <- as.Date(paste(year(y),i,'01',sep='-'))
      if (j==1) Y <- zoo(y) else
                Y <- c(Y,zoo(y))
      j <- j+1
    }
  }
  Y <- attrcp(X,Y)
  class(Y) <- class(X)
  attr(Y,'dimnames') <- NULL
  invisible(Y)
}

@@

```{r compare}
comparemannual <- function(tas_DECA,reanalysis='~/R/air.mon.mean.nc',mask=TRUE,l= 3,eofs=NULL) {

@@

```{r diagnose}
diagnoseMOS <- function(results,y,l=1) {
  print(paste('diagnoseMOS',results))
  load(results)  
  zl <- regexpr('.z.',results)
  print(paste('zmos <-',substr(results,zl+1,nchar(results)-4)))
  eval(parse(text=paste('zmos <-',substr(results,zl+1,nchar(results)-4))))
  attr(zmos,'variable') <- varid(y)
  print(range(year(zmos)))
  y <- pentad(y,it0=min(year(zmos)),l=l)
  print(index(y))
  print(index(zmos))
  print('ensemble mean: aggregate(.)')
  ## Different ensemble emmbers are represented as calendar months
  Z <- aggregate(zmos,year,'mean')
  index(Z) <- year(Z)
  print('correlation: ensemble mean FCST & reanalysis')
  r <- corfield(Z,y,plot=FALSE)
  map(r,colbar=list(breaks=seq(-1,1,by=0.1),pal='t2m',rev=TRUE),type='fill',new=FALSE)
  figlab('EOBS+ensemble mean MOS',ypos=0.999)

@@

print('Get the predictands: mu & fw from EOBS')
if (!file.exists('~/storeA/mu_0.25deg_reg.nc'))
  mu.eobs <-  wetdaymean('~/data/data.ECAD/rr_0.25deg_reg.nc') else
  mu.eobs <- retrieve('~/storeA/mu_0.25deg_reg.nc')
fw.eobs <- retrieve('~/storeA/fw_0.25deg_reg.nc')
attr(mu.eobs,'variable') <- 'mu'
attr(fw.eobs,'variable') <- 'fw'
attr(fw.eobs,'unit') <- 'fraction'
#index(fw.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))
#index(mu.eobs) <- year(as.Date(time,origin = as.Date("1950-01-01")))

@@

## Test to examine the link between large-scale temperature and wet-day mean precip
## Derive the saturation vapour pressure from temperature
## retrieve did not get the right time stamp:
  es <- annual(C.C.eq(retrieve('~/R/air.mon.mean.nc',lon=c(-60,40),lat=c(20,70))))
  es.eof <- EOF(mask(es,land=TRUE))

## Also get the mean sea-level pressure for the wet-day frequency 
  slp <- annual(retrieve('~/R/slp.mon.mean.nc',lon=c(-60,40),lat=c(20,70)))
  slp.eof <- EOF(slp)

@@

  class(mu.eof) <- class(es.eof)
  index(mu.eof) <- year(mu.eof)
  mu.ds <- DS(mu.eof,es.eof)
  class(fw.eof) <- class(slp.eof)
  index(fw.eof) <- year(fw.eof)
  fw.ds <- DS(fw.eof,slp.eof)

  plot(mu.ds,colbar1=list(breaks=seq(-0.025,0.025,by=0.005),rev=TRUE),
       colbar2=list(breaks=seq(-350,350,by=25)),type='fill',new=FALSE)
  figlab('ESD(NCEP/NCAR)',ypos=0.999)

@@

print('Get the decadal predictors saved locally in "~/data/gfdl_DECA.rda"')
load('~/R/mme_DECA.rda')
```

@@

if (looexperiments) {
## Carry out a LOO cross-validation experiment
  print('Check results for LOO experiment - if absent do it')
  if (!file.exists('~/R/specsdecadal.z.fw.y1.rda')) {
    print('do specsdecadal.z.fw.y1.rda')
    z.fw.y1 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1,it=NULL,is=NULL,mask=FALSE)
    save(file='~/R/specsdecadal.z.fw.y1.rda',z.fw.y1)
    rm("z.fw.y1"); gc(reset=TRUE)
  }
  if (!file.exists('~/R/specsdecadal.z.mu.y1.rda')) {
    print('do specsdecadal.z.mu.y1.rda')
    z.mu.y1 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1,it=NULL,is=NULL,mask=TRUE)
    save(file='~/R/specsdecadal.z.mu.y1.rda',z.mu.y1)
    rm("z.mu.y1"); gc(reset=TRUE)
  }
  if (!file.exists('~/R/specsdecadal.z.fw.y1.3.rda')) {
    print('do specsdecadal.z.fw.y1.3.rda')
    z.fw.y1.3 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:3,it=NULL,is=NULL,mask=FALSE)
    save(file='~/R/specsdecadal.z.fw.y1.3.rda',z.fw.y1.3)
    rm("z.fw.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists('~/R/specsdecadal.z.mu.y1.3.rda')) {
    print('do specsdecadal.z.mu.y1.3.rda')
    z.mu.y1.3 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:3,it=NULL,is=NULL,mask=TRUE)
    save(file='~/R/specsdecadal.z.mu.y1.3.rda',z.mu.y1.3)
    rm("z.mu.y1.3"); gc(reset=TRUE)
  }
  if (!file.exists('~/R/specsdecadal.z.fw.y1.5.rda')) {
    print('do specsdecadal.z.fw.y1.5.rda')
    z.fw.y1.5 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:5,it=NULL,is=NULL,mask=FALSE)
    save(file='~/R/specsdecadal.z.fw.y1.5.rda',z.fw.y1.5)
    rm("z.fw.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists('~/R/specsdecadal.z.mu.y1.5.rda')) {
    print('do specsdecadal.z.mu.y1.5.rda')
    z.mu.y1.5 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:5,it=NULL,is=NULL,mask=TRUE)
    save(file='~/R/specsdecadal.z.mu.y1.5.rda',z.mu.y1.5)
    rm("z.mu.y1.5"); gc(reset=TRUE)
  }
  if (!file.exists('~/R/specsdecadal.z.fw.y1.9.rda')) {
    print('do specsdecadal.z.fw.y1.9.rda')
     z.fw.y1.9 <- D53Exp(X=psl_DECA,Y=fw.eobs,m=1:9,it=NULL,is=NULL,mask=FALSE)
     save(file='~/R/specsdecadal.z.fw.y1.9.rda',z.fw.y1.9)
     rm("z.fw.y1.9"); gc(reset=TRUE)
   }
  if (!file.exists('~/R/specsdecadal.z.mu.y1.9.rda')) {
    print('do specsdecadal.z.mu.y1.9.rda')
    z.mu.y1.9 <- D53Exp(X=tas_DECA,Y=mu.eobs,m=1:9,it=NULL,is=NULL,mask=TRUE)
    save(file='~/R/specsdecadal.z.mu.y1.9.rda',z.mu.y1.9)
    rm("z.mu.y1.9"); gc(reset=TRUE)
  }

## Diagnostics:
  print('diagnose the MOS results')
  load('~/R/specsdecadal.z.mu.y1.3.rda')
  it0 <- range(year(z.mu.y1.3))
  print(paste('time period',paste(it0,collapse='-')))
  score.mu.y1 <- diagnoseMOS('~/R/specsdecadal.z.mu.y1.rda',mu.eobs)
  score.fw.y1 <- diagnoseMOS('~/R/specsdecadal.z.fw.y1.rda',fw.eobs)
  score.mu.y1.3 <- diagnoseMOS('~/R/specsdecadal.z.mu.y1.3.rda',mu.eobs,l=3)
  score.fw.y1.3 <- diagnoseMOS('~/R/specsdecadal.z.fw.y1.3.rda',fw.eobs,l=3)
  score.mu.y1.5 <- diagnoseMOS('~/R/specsdecadal.z.mu.y1.5.rda',mu.eobs,l=5)
  score.fw.y1.5 <- diagnoseMOS('~/R/specsdecadal.z.fw.y1.5.rda',fw.eobs,l=5)
  score.mu.y1.9 <- diagnoseMOS('~/R/specsdecadal.z.mu.y1.9.rda',mu.eobs,l=9)
  score.fw.y1.9 <- diagnoseMOS('~/R/specsdecadal.z.fw.y1.9.rda',fw.eobs,l=9)
  while (dev.cur()>1) dev.off()
## Repeat and make a table of the results...

@@

if (diagnoseFCST) {
  comparemannual(tas_DECA,mask=TRUE,l= 1)
  comparemannual(tas_DECA,mask=TRUE,l= 2)
  comparemannual(tas_DECA,mask=TRUE,l= 4)
  comparemannual(tas_DECA,mask=TRUE,l= 5)
  comparemannual(tas_DECA,mask=TRUE,l= 6)
  comparemannual(tas_DECA,mask=TRUE,l= 7)
  comparemannual(tas_DECA,mask=TRUE,l= 8)
  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 1)
  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 2)
  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 3)
  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 4)
  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 5)
  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 6)
  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 7)
  comparemannual(psl_DECA,reanalysis='~/R/slp.mon.mean.nc',mask=FALSE,l= 8)
  rmap.1.10.mask.3 <- comparemannual(tas_DECA,mask=TRUE,l= 3, eofs=1:10)
  rmap.1.10.nomask.3 <- comparemannual(tas_DECA,mask=FALSE,l= 3, eofs=1:10)
}
```

@@

```{r t2m scales}
## Scaling dependency: time - space
t2m <- retrieve('~/R/air.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
t2m <- anomaly(t2m)
zst <- scaledependency(t2m)
#zst[,1] <- NA
d <- dim(zst)

@@

## Repeat for SLP
slp <- retrieve('~/R/slp.mon.mean.nc',lon=c(-90,60),lat=c(-20,80))
slp <- anomaly(slp)
yst <- scaledependency(slp)
image(attr(yst,'timescales'),attr(yst,'gridboxes'),t(t(yst)/colMeans(yst)),
      main='Mean sea-level pressure: spectral power',
      xlab='Temporal scale (years)',ylab='spatial scale (grid boxes)')
contour(attr(yst,'timescales'),attr(yst,'gridboxes'),yst,add=TRUE)
